### 2017-03-19 Manuscript/SW_early_life_telomeres.Rmd
Home edits

Each time a bird is caught on Cousin a range of morphometric measurements are taken, including body mass and tarsus length (to the nearest 0.1g and 0.1mm, respectively). A blood sample (*ca* 25 l) is taken via brachial venipuncture, and stored at room temperature in 1 ml of absolute ethanol in a 1.5 ml screw-cap microfuge tube.

@@

We used two approaches to determine individual-level consistency in RTL. We first calculated individual-level repeatability in RTL by dividing the random avariance explained by individual ID by the total random variance, in a model of that accounted for age and cohort effects. Second, we constructed a LMM with RTL at time *t* + 1 as the response variable, RTL at time *t* and age at time *t* as fixed effects, and individual ID and cohort as random effects. We estimated the slope of the relationship between within-individual telomere measurements, as well as the variance explained, by calculating the marginal R^2^ [@Nakagawa2013] of the model.

@@

We also tested whether consistent telomere lengthening acrossed out dataset using a modified version of the approach recently developed by Simons *et al.* [-@Simons2014]. Briefly, this approach utilises samples with at least three telomere measurements to compare residual variance in telomere change over time with the overall change in telomere length between the first and least telomere measurements [@Simons2014]. If, in samples that exhibit lengthening, the overall increase in telomere length exceeds the residual variance, then telomere lengthening is predicted to be occurring [@Simons2014]. If, on the other hand, increases in telomere length are due to measurement error, within-individual residual variance in telomere length is expected to be similar to overall observed increases in telomere length.

@@

#####Telomere dynamics and age among cohorts
We first tested how RTL was related to age among cohorts using a model selection approach. The top model contained a log-linear relationship between RTL and age (Table 1A). All other models fitted the data much less well ($\Delta$AICc > 15; Table 1). The log-linear relationship between RTL and age could be seen clearly in the raw data; RTL decreased with age (estimate = `r round3(fixef(RTL_LogAge)['LogAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_LogAge)['LogAge',],lessthan = F),collapse = ', ')`), with the greatest decrease occurring in the first year of life (Fig. 1A). There was substantial variation in RTL among cohorts, with no obvious trend over time (Fig. 1B). There was a negative relationship between RTL and age in 21 of the `r length(table(dd$Cohort))` cohorts, but the slope of the relationship between RTL and log age varied substantially among cohorts (Fig. 1C). To test whether this variation was significant we fitted a model including the log age x cohort interaction term, and found that this was a marginally better fit than a model including only main effects ($\Delta$AICc = `r round2(AICc(RTL_LogAge) - AICc(RTL_LogAgeXcohort))`). In the one year in which RTL increased with age (2013), 17 of the 18 birds sampled were fledglings or subadults, suggesting that the observed pattern was an artefact of the sampling in this season (i.e. a lack of variation in age among sampled birds), rather than a real relationship.

@@

Using the approach outlined by Simons *et al.* we tested whether, statistically, overall increases in RTL over lifespans could be detected statistically in our dataset. We found no evidence that this was the case - overall increases in RTL within individuals did not exceed residual variance. In fact, the opposite was the case - residual variance in RTL was significantly greater than observed RTL increases over lifespans (P = 0.02). This suggests that increases in RTL within individuals are sporadic, and not consistent over individual lifespans.

>>>>

Each time a bird is caught on Cousin a range of morphometric measurements are taken, including body mass and tarsus length (to the nearest 0.1g and 0.1mm, respectively). A blood sample (*ca* 25 $\mu$l) is taken via brachial venipuncture, and stored at room temperature in 1 ml of absolute ethanol in a 1.5 ml screw-cap microfuge tube.

@@

We used two approaches to determine individual-level consistency in RTL. We first calculated individual-level repeatability in RTL by dividing the random variance explained by individual ID by the total random variance, in a model of that accounted for age and cohort effects. Second, we constructed a LMM with RTL at time *t* + 1 as the response variable, RTL at time *t* and age at time *t* as fixed effects, and individual ID and cohort as random effects. We estimated the slope of the relationship between within-individual telomere measurements, as well as the variance explained, by calculating the marginal R^2^ [@Nakagawa2013] of the model.

@@

We also tested whether consistent telomere lengthening across our dataset using a modified version of the approach developed by Simons *et al.* [-@Simons2014]. Briefly, this approach utilises samples with at least three telomere measurements to compare residual variance in telomere change over time with the overall change in telomere length between the first and least telomere measurements [@Simons2014]. If, in samples that increase in length, the overall increase in telomere length exceeds the residual variance, then telomere lengthening cannot be explained by error [@Simons2014]. If, on the other hand, increases in telomere length are due to measurement error, within-individual residual variance in telomere length is expected to be similar to overall observed increases in telomere length.

@@

#####Telomere dynamics and age among cohorts
We first tested how RTL was related to age among cohorts using a model selection approach. The top model contained a log-linear relationship between RTL and age (Table 1A). All other models fitted the data much less well ($\Delta$AICc > 15; Table 1). The log-linear relationship between RTL and age could be seen clearly in the raw data; RTL decreased with age (estimate = `r round3(fixef(RTL_LogAge)['LogAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_LogAge)['LogAge',],lessthan = F),collapse = ', ')`), with the greatest decrease occurring in the first year of life (Fig. 1A). There was substantial variation in RTL among cohorts, with no obvious trend over time (Fig. 1B). There was a negative relationship between RTL and log age in 21 of the `r length(table(dd$Cohort))` cohorts, but the slope the relationship varied substantially among cohorts (Fig. 1C). To test whether this variation was significant we fitted a model including the log age x cohort interaction term, and found that this was a marginally better fit than a model including only main effects ($\Delta$AICc = `r round2(AICc(RTL_LogAge) - AICc(RTL_LogAgeXcohort))`). In the one year in which RTL increased with age (2013), 17 of the 18 birds sampled were fledglings or subadults, suggesting that the observed pattern was an artefact of the sampling in this season (i.e. a lack of variation in age among sampled birds), rather than a real relationship.

@@

Using the approach outlined by Simons *et al.* we tested whether overall increases in RTL over lifespans could be detected statistically in our dataset. We found no evidence that this was the case: overall increases in RTL within individuals did not exceed residual variance; in fact, residual variance in RTL was significantly greater than observed RTL increases over lifespans (P = 0.02). This suggests that increases in RTL within individuals are sporadic, and not consistent over individual lifespans.
